On hyperparameter tuning through Lipschitz global optimization

### 2024-07-25 2024/ITMO paper/sn-article.tex
fix first reviewer comments

\begin{remark}
The Lipshitz condition will not be satisfied for the reduced one-dimensional function $\varphi(y(x))$. However, many one-dimensional Lipschitz optimization algorithms can be generalized to the case of H{\"o}lder function minimization. A concrete example of such a generalization is given in subsection \ref{sec_GSA}.

@@

@@

Let us now consider the mapping
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
which translates any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
\begin{equation}\label{f}
f(x) = \varphi(Y(x),u_{E(x)+1}), x\in[0,S],
\end{equation}
which has, generally speaking, discontinuities at integer points $x_i = i, 1 \leq i \leq S-1$.
Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
%, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.

@@

The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods. The application area of the framework is focused on the hyperparameter tuning tasks, since this class of tasks involves the problems of multiextremality of the optimization criterion and the lack of a formula-based description of the model under study (a ``black box'' type model).

>>>>

\begin{remark}
The Lipschitz condition will not be satisfied for the reduced one-dimensional function $\varphi(y(x))$. However, many one-dimensional Lipschitz optimization algorithms can be generalized to the case of H{\"o}lder function minimization. A concrete example of such a generalization is given in subsection \ref{sec_GSA}.

@@

\textcolor[rgb]{0,0,1}{
\begin{remark}
In this work we use piecewise-linear approximation $y_m(x)$ to the true curve $y(x)$ (Fig.~\ref{fig:Peano}). It is simple from the point of view of implementation and use, but does not reflect all the properties of the Peano-Hilbert curve. More complex approximations fit the curve better. For example, in \cite{Strongin2000} a non-univalent approximation was also proposed. It reflects the following property of the Peano-Hilbert curve $y(x)$: a point in $D$ can have several (but not more than $2^N$) inverse images in $[0,1]$. Using non-univalent approximation allows one to reduce the number of objective function evaluations required to solve the problem with a given accuracy (see \cite{Sergeyev2022, Lera2024}).
\end{remark}}

@@

Let us now consider the mapping
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
which maps any point of the interval $[0,S]$ to the region $D$ (the notation $E(x)$ corresponds to the integer part of the number $x$) and define the function
\begin{equation}\label{f}
f(x) = \varphi(Y(x),u_{E(x)+1}), x\in[0,S],
\end{equation}
which has, generally speaking, discontinuities at integer points $x_i = i, 1 \leq i \leq S-1$.
Therefore, the values $z_i = f(x_i)$ at these points will be considered undefined and will not be used in the algorithm.
%, а значения индекса -- равным 0, т.е. $\nu(x_i) = 0$.

@@

The iOpt framework can be used to develop specialized decision support systems for solving the tasks of selecting optimal parameters for complex models or methods.
\textcolor[rgb]{0,0,1}{In particular, it is focused on solving the problems of tuning hyperparameters of machine learning methods. In such problems there is a lack of analytical description of the objective function (``black box'' type problems), and the function itself can be multi-extremal.}
